---
title: "Lab 08 - University of Edinburgh Art Collection"
author: "Noah Booker"
date: "4/4/25"
output: github_document
---

## Load Packages and Data  

First, let's load the necessary packages:  

```{r load-packages, message = FALSE}
library(tidyverse) 
library(skimr)
```

Checking whether bots are allowed to access pages on this domain.

```{r}
library(robotstxt)
paths_allowed("https://collections.ed.ac.uk/art)")
```

Now, load the dataset. If your data isn't ready yet, you can leave `eval = FALSE` for now and update it when needed.  


```{r load-data, message = FALSE}
uoe_art <- read_csv("data/uoe-art.csv")
```

## Exercise 9

Let's start working with the **title** column by separating the title and the date:  

```{r separate-title-date, error = TRUE}
uoe_art <- uoe_art %>%
  separate(title, into = c("title", "date"), sep = "\\(") %>%
  mutate(year = str_remove(date, "\\)") %>% as.numeric()) %>%
  select(title, artist, year, link)  # Fill in the missing variable!
```

## Exercise 10  

Note that this approach will result in some warnings when you run the code, and that’s OK! Read the warnings, and explain what they mean, and why we are ok with leaving them in given that our objective is to just capture year where it’s convenient to do so.

It seems that the first warning is referring to titles that have parentheticals before the date, such as piece number 2, A World under Your Feet/ Composition with Yellow Crosses (diptych) (1990), and piece number 85, Frieze from the Temple of Athena Nike (left slab) (1827). In those cases, the code above does not serve our purposes—it removes an informative element of the title and does not extract the date.

The second warning, I think, is warning us that the mutate function is returning some NA values. For the pieces with titles without dates, this is fine. However, as explained above, I think this also includes some titles with dates as well.

## Exercise 11  

Print out a summary of the data frame using the skim() function. How many pieces have artist info missing? How many have year info missing?

```{r skim}
skim(uoe_art)
```
There are 111 pieces without artist info and 1,575 pieces without year info.

## Exercise 12  

Make a histogram of years. Use a reasonable bin width. Do you see anything out of the ordinary?

```{r year_plot}
uoe_art %>% 
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 10)
```

There seems to be at least one data point near year 0. 


## Exercise 13

Find which piece has the out-of-the-ordinary year and go to its page on the art collection website to find the correct year for it. Can you tell why our code didn’t capture the correct year information? Correct the error in the data frame and visualize the data again.

Piece number 1871 originally had the title Death Mask (2) (1964). Our code to extract the year instead extracted 2 because it's the numeric value that appeared after the first open parentheses.

```{r fix_year}
uoe_art <- uoe_art %>%
  mutate(year = if_else(row_number() == 1871, 1964, year))

uoe_art %>% 
  ggplot(aes(x = year)) +
  geom_histogram(binwidth = 1)
```

## Exercise 14

Who is the most commonly featured artist in the collection? Do you know them? Any guess as to why the university has so many pieces from them?

```{r modal_artist}
uoe_art %>% 
  count(artist) %>%
  arrange(desc(n)) %>% 
  head()
```

The most common artist is Emma Gillies. Wikipedia says that "Emma Smith Gillies (1900–1936) was a Scottish potter best known for her early adaptation of Art Deco painted vases and jugs." It also says that was born near Edinburhg, attended pottery classes at Edinburgh College of Art, studying there under a renowned sculptor, and that some of her work was discovered (posthumously) at University of Edinburgh in 2012.

